FraCaS: Temporal Analysis

### 2020-10-01 paper2020/paper.tex
repair citation
kill todo and minor fixes

\usepackage[hyperref]{naaclhlt2019}
\usepackage{natbib}
\usepackage{tabularx}
\usepackage{times}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{todonotes}
\usepackage{enumitem}
\usepackage{cleveref}
\newlist{lingex}{enumerate}{3}
\setlist[lingex,1]{parsep=0pt,itemsep=1pt,label=(\arabic*),resume=lingexcount}
\newcommand\onelingex[1]{\begin{lingex}\item #1 \end{lingex}}

@@

The semantics of tense and aspect has been a long standing issue in
the study of formal semantics since the early days of Montague Grammar
and a number of different ideas have been put forth on the table to
deal with them throughout the years. Recent proposals include the
works of the following authors:
\citet{dowty:2012,prior:2003,steedman_productions_2000,higginbotham:2009,fernando:2015}.
%
The
semantics of tense and aspect have been also considered in the study of
Natural Language Inference (NLI). The various datasets
for NLI that have been proposed by the years contain examples that
have some implicit or explicit reliance on inferences related to tense
and aspect. One of the early datasets used to test logical approaches,
the FraCaS test suite~\citep{cooper:1996} contains a whole section
dedicated to temporal and aspectual inference (section 7 of the dataset). This part
of the FraCaS test suite has been difficult to tackle. That is,
so far, no computational system has been capable to deal with it: when
authors report accuracy over the FraCaS test suite they
skip this section.
%
In fact, they also often skip the anaphora and ellipsis sections, the
exception being the system presented by
\citet{bernardy_type-theoretical_2017,bernardy_wide-coverage_2019},
which include support for anaphora and ellipsis but still omit the
temporal section.
%
In this paper, we take up the challenge of providing a
computationally viable account of tense and aspect to deal with the
notoriously difficult section 7 of the FraCaS test suite. Our account is not meant
to be a theoretically extensive account of tense and aspect, but rather an
account that is driven by the need to cover the test suite in a way
that is general enough to capture the test suite examples, \emph{while still covering the rest of the FraCaS test suite}.
%
The account is evaluated on the entailment properties of various
temporal and aspectual examples, as given by the test suite. As such,
we are not getting into the discussion of how tense and aspect might
affect grammaticality or infelicitousness of various sentences.
We assume that the sentences of the FraCaS suite are syntactically and
semantically correct, and strive to produce accurate logical
representations given that assumption. We further assume that the
entailment annotations of various problems are valid, and we use those
to evaluate the correctness of the logical representations of
sentences.

@@

For the purpose of this paper, the important point is that the GF
syntax aims at a balance of sufficient abstraction to provide a
semantically-relevant structure, but at the same time it embeds
sufficiently many syntactic features to be able to reconstruct
natural-language text. That is, the parse trees generally satisfy the
homomorphism requirement of \citet{montague_english_1970,montague_proper_1974}, and
we can focus on the translation of syntactic trees to logical forms.
Example \ref{ex:nobel} shows an example abstract syntax tree and its
realization in English.

\paragraph{Coq}
Coq is an interactive theorem prover (proof assistant). It is based on
the calculus of co-inductive constructions~\citep{werner_une_1994} We
do not use any co-induction (or even induction) in this paper, relying
on the pure lambda-calculus inner core of Coq.  Coq is a very powerful
reasoning engine that makes it fit for implementing natural language
semantics.  Coq supports, as already mentioned, dependent typing, but
also subtyping. Both concepts are instrumental in expressing NL
semantics\citet{TODO: chatzikyriakidis and luo}.  Besides, on a more
practical side, it works well for the the task of NLI, when the latter
is formalized as a theorem proving task: its many tactics mean that
many tasks in theorem proving are trivialized. In particular, all
problems of time-intervals inclusion, which occur in every temporal
problems, are solved with Coq's linear arithmetic tactic.

>>>>

\usepackage[hyperref]{naaclhlt2019}
\usepackage{natbib}
\usepackage{tabularx}
\usepackage{times}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{todonotes}
\usepackage{enumitem}
\usepackage[capitalize]{cleveref}
\newlist{lingex}{enumerate}{3}
\setlist[lingex,1]{parsep=0pt,itemsep=1pt,label=(\arabic*),resume=lingexcount}
\newcommand\onelingex[1]{\begin{lingex}\item #1 \end{lingex}}

@@

The semantics of tense and aspect has been a long standing issue in
the study of formal semantics since the early days of Montague Grammar
and a number of different ideas have been put forth on the table to
deal with them throughout the years. Recent proposals include the
works of the following authors:
\citet{dowty:2012,prior:2003,steedman_productions_2000,higginbotham:2009,fernando:2015}.
%
The
semantics of tense and aspect have been also considered in the study of
Natural Language Inference (NLI). The various datasets
for NLI that have been proposed by the years contain examples that
have some implicit or explicit reliance on inferences related to tense
and aspect. One of the early datasets used to test logical approaches,
the FraCaS test suite~\citep{cooper:1996} contains a whole section
dedicated to temporal and aspectual inference (section 7 of the dataset). This part
of the FraCaS test suite has been difficult to tackle. That is,
so far, no computational system has been capable to deal with it: when
authors report accuracy over the FraCaS test suite they
skip this section.
%
In fact, they also often skip the anaphora and ellipsis sections, the
exception being the system presented by
\citet{bernardy_type-theoretical_2017,bernardy_wide-coverage_2019},
which includes support for anaphora and ellipsis but still omit the
temporal section.
%
In this paper, we take up the challenge of providing a
computationally viable account of tense and aspect to deal with the
notoriously difficult section 7 of the FraCaS test suite. Our account is not meant
to be a theoretically extensive account of tense and aspect, but rather an
account that is driven by the need to cover the test suite in a way
that is general enough to capture the test suite examples, \emph{while still covering the rest of the FraCaS test suite}.
%
The account is evaluated on the entailment properties of various
temporal and aspectual examples, as given by the test suite. As such,
we are not getting into the discussion of how tense and aspect might
affect grammaticality or infelicitousness of various sentences.
We assume that the sentences of the FraCaS suite are syntactically and
semantically correct, and strive to produce accurate logical
representations given that assumption. We further assume that the
entailment annotations of various problems are valid, and we use those
to evaluate the correctness of the logical representations of
sentences.

@@

For the purpose of this paper, the important feature of GF
syntax is that it aims at a balance of sufficient abstraction to provide a
semantically-relevant structure, but at the same time it embeds
sufficiently many syntactic features to be able to reconstruct
natural-language text. That is, the parse trees generally satisfy the
homomorphism requirement of \citet{montague_english_1970,montague_proper_1974}, and
we can focus on the translation of syntactic trees to logical forms.
%
Consequently, the system presented
here does not aim at textual natural language understanding, but rather
provides a testable, systematic formal semantics of temporal phenomena.
Example \ref{ex:nobel} shows an example abstract syntax tree and its
realization in English.

\paragraph{Coq}
Coq is an interactive theorem prover (proof assistant). It is based on
the calculus of co-inductive constructions~\citep{werner_une_1994} We
do not use any co-induction (or even induction) in this paper, relying
on the pure lambda-calculus inner core of Coq.  Coq is a very powerful
reasoning engine that makes it fit for implementing natural language
semantics.  Coq supports, as already mentioned, dependent typing, but
also subtyping. Both concepts are instrumental in expressing NL
semantics\citet{chatzikyriakidis_natural_2014}.  Besides, on a more
practical side, it works well for the the task of NLI, when the latter
is formalized as a theorem proving task: its many tactics mean that
many tasks in theorem proving are trivialized. In particular, all
problems of time-intervals inclusion, which occur in every temporal
problems, are solved with Coq's linear arithmetic tactic.
